They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-21 paper/paper.tex
adding breakdown by round i

\hypertarget{description-of-call-of-duty-search-and-destroy}{%
\subsection{Description of Call of Duty Search and
Destroy}\label{description-of-call-of-duty-search-and-destroy}}

@@

The empirical offensive round win percentage across all rounds is
\(\phi_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global average.

@@

\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(r_+) = {\binom {r}{r_+}} \omega^{r_+}(1-\omega)^{r-\sigma}.
}\label{eq:binom}\end{equation}

@@

Stepping back from the one-sided perspective of a single team's
probability of winning and their streakiness, I can attempt to detect
the ``hot hand'' more broadly with the Wald-Wolfowitz runs test, a
common tool used in the field of quality control. Under the null
hypothesis, the number of runs in a sequence of \(r\) trials is a random
variable with the following mean \(\mu\) and variance \(\sigma^2\).

@@

@@

@@

For the sake of brevity, I did not delve into win rates and streaks
split by offensive or defensive role. For example, it would be
interesting to look at whether the round win rate is higher after a
streaks of three round wins when the team starts the streak winning as
the offensive team. This would mean that, in the round immediately
following the streak, the streaking team would be playing defense, where
teams are slightly more likely to win on average. On the other hand,
streaks of three starting with an offensive round win---theoretically,
given \(\phi_O\) = 47.7\% the expected frequency is
\(\phi_O^2 \times (1 - \phi_O) \approx 0.119\)---are likely to occur
less frequently than such streaks starting with a defensive
win---\(\phi_O \times (1 - \phi_O)^2 \approx 0.130\)---so I would want
to account for sample size differences.

Regarding future work, I've reached out to the Twitter user
``R11stats'', who privately tracks in-round player
engagements.\footnote{\url{https://twitter.com/R11stats}} R11stats
expressed intent on making the data public, which would allow for
research into player-specific momentum.

@@

I would like to thank my brother Andrew for giving me feedback on early
iterations of this paper, and for requiring me to mention him.

@@

>>>>

\hypertarget{game-description}{%
\subsection{Game Description}\label{game-description}}

@@

The empirical offensive round win percentage across all rounds is
\(\tau_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global
average.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense or defense in either the
  SnD series played as the second game or the fifth game in the
  best-of-five format used in tournaments, and their opponent is
  assigned to play offense in the first round of the SnD slot (second or
  fifth game) not chosen. ({``Competitive Settings \textbar{} Call of
  Duty League,''} n.d.) (Other game variants like Hardpoint are played
  as the first, third, and fourth games in the matchup.) Anecdotally,
  the higher-seeded teams tend to choose to play defense in the first
  round of the SnD series played as the second game, although this
  choice is not universal. With online qualifiers, teams all play each
  other twice throughout the season. Each team gets to play the
  higher-seeded role once in their matchups, despite their win-loss
  records.}

@@

\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(r_+) = {\binom {r}{r_+}} \omega^{r_+}(1-\omega)^{r-r+}.
}\label{eq:binom}\end{equation}

@@

Stepping back from the one-sided perspective of a single team's
probability of winning and their streakiness, one can attempt to detect
the ``hot hand'' more broadly with the Wald-Wolfowitz runs test, a
common tool used in the field of quality control. Under the null
hypothesis, the number of runs in a sequence of \(r\) trials is a random
variable with the following mean \(\mu\) and variance \(\sigma^2\).

@@

One might subset observed series sequences to those that violate the run
test null hypothesis and perform a proportions test where the null is
that the observe rate of such series is equal to the frequency that we
should expect, given a choice for the constant round win probability
\(\phi\). If the

@@

Now let us explicitly consider the round, \(i\). As shown in Table
\ref{tbl:cod-pw3ri-pl3ri}, the notional win rate in the round \(i\)
immediately following a streak of \(k\) round wins in a series lasting
\(r\) rounds, \(\hat{P}{w|kri}_0\), is statistically different than the
observed \(P^{w|kri}\) in several cases, notably when \(r = i\) for each
of \(r \in [9, 10, 11]\).\footnote{The null can also be rejected for
  \(r \in [7, 8]\), although these are not shown, and the sample sizes
  are smaller.} Interestingly, the observed rate is greater than the
expected rate in each case where we can reject the null hypothesis,
whereas the observed rate is less than the expected rate in all other
cases (as well in a strong majority of cases not shown).

\begin{longtable}{rrrrrr}
\caption{Given the round $i$, the round win streak $k=3$, the length of the series ($r$ rounds), and the series winner, the observed count and proportion of round wins immediately following the streak, $N^{w|kri}$ and $P^{w|kri}$ respectively. Additionally, the notional proportion, $\hat{P}^{w|kri}_0$, and the $p$-value of a binomial test comparing the underlying notional $p^{w|kri}_0$ to the probability implied by the observed proportion $p^{w|kri} = N^{w|kri} \times P^{w|kri}$. Table restricted to $r \in [9, 10, 11]$ and $r - 2 \leq i \leq r$ for brevity.}
\label{tbl:cod-pw3ri-pl3ri} \\
\toprule
$r$ & $i$ & \text{Win series?} & $N^{w|kri}$ & $P^{w|kri}$ & $\hat{P}^{w|kri}_0$ ($p$-value) & $\hat{P}^{w|kri}_0 > P^{w|kri}$ \\

9 & 7 & yes & 36 & 41.7\% & 50.0\% (0.41) & no \\ 
9 & 8 & yes & 40 & 45.0\% & 50.0\% (0.64) & no \\ 
9 & 9 & yes & 37 & 89.2\% & 50.0\% (<=0.01) & yes \\ 
10 & 8 & no & 37 & 2.7\% & 14.3\% (0.06) & no \\ 
10 & 8 & yes & 37 & 24.3\% & 42.9\% (0.03) & no \\ 
10 & 9 & no & 32 & 6.2\% & 14.3\% (0.31) & no \\ 
10 & 9 & yes & 32 & 12.5\% & 42.9\% (<=0.01) & no \\ 
10 & 10 & yes & 29 & 72.4\% & 42.9\% (<=0.01) & yes \\ 
11 & 9 & no & 32 & 15.6\% & 25.0\% (0.31) & no \\ 
11 & 9 & yes & 32 & 31.2\% & 37.5\% (0.58) & no \\ 
11 & 10 & no & 36 & 13.9\% & 25.0\% (0.18) & no \\ 
11 & 10 & yes & 36 & 13.9\% & 37.5\% (<=0.01) & no \\ 
11 & 11 & yes & 32 & 53.1\% & 37.5\% (0.10) & yes \\

@@

For the sake of brevity, I did not delve into win rates and streaks
split by offensive or defensive role. For example, it would be
interesting to look at whether the round win rate is higher after a
streaks of three round wins when the team starts the streak winning as
the offensive team. This would mean that, in the round immediately
following the streak, the streaking team would be playing defense, where
teams are slightly more likely to win on average. On the other hand,
streaks of three starting with an offensive round win---theoretically,
given \(\tau_O\) = 47.7\% the expected frequency is
\(\tau_O^2 \times (1 - \tau_O) \approx 0.119\)---are likely to occur
less frequently than such streaks starting with a defensive
win---\(\tau_O \times (1 - \tau_O)^2 \approx 0.130\)---so I would want
to account for sample size differences.

Further regarding future work, I've reached out to the Twitter user
``R11stats'', who privately tracks in-round player
engagements.\footnote{\url{https://twitter.com/R11stats}} R11stats
expressed intent on making the data public, which would allow for
research into player-specific momentum.

@@

I would like to thank my brother Andrew for giving me feedback on early
iterations of this paper.

@@

\leavevmode\vadjust pre{\hypertarget{ref-codrules}{}}%
{``Competitive Settings \textbar{} Call of Duty League.''} n.d.
\url{https://callofdutyleague.com/en-us/competitive-settings?tab=map-veto}.
